---
title: "BUAN6357-HW4B"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
date: "2023-02-20"
editor_options: 
  markdown: 
    wrap: 72
---

Adding libraries and datasets

```{r}
rm(list=ls())
demo <- T
data("airquality")
require(data.table)
require(tidyverse)
```

Declaring seed and cleaning the working dataset 

```{r}
seed    <- 504737137
raw     <- setDT(na.omit(airquality))[,-c("Day")]
raw$Month <- as.factor(raw$Month)
(n      <- nrow(raw))
```

Linear model relation

```{r}
fmla <- Ozone ~ .
```

Baseline model

```{r}
base.resid <- lm(fmla,data=raw)$residuals

if(demo){summary(lm(fmla,data=raw))}
```

Simple cross validation model

```{r}

set.seed(seed)
frac <- 0.1
tst <- sample(n,ceiling(frac*n))
m1  <- lm(fmla,data=raw[-tst,])
if(demo){summary(m1)}
cv.trn <- m1$fitted.values
if(demo){summary(cv.trn)}
cv.tst <- predict(m1,raw[tst,])
if(demo){summary(cv.tst)}
cv.resid <- data.table(loc=tst,diff=raw$Ozone[tst]-cv.tst)
if(demo){summary(cv.resid)}

```


Function to run linear model on training and testing data sets

```{r}

fitPred <- function(trn,tst,i) {
  m <- lm(fmla,data=trn)  
  y <- tst$Ozone    
  p <- predict(m,tst)     
  r <- y - p              
  return(data.table(idx=i,resid=r,y=y,yhat=p))
}

```

Jackknife model LOOCV

```{r}

c2      <- data.table(grp=1:n, tst=1:n)
dt2     <- c2[,fitPred(raw[-tst,],raw[tst,],tst),by=.(grp)]
jk.resid      <- data.table(loc=dt2$idx,diff=dt2$resid)
if(demo){summary(jk.resid)}

```


K fold model


```{r}

set.seed(seed)
kf  <- 10
t       <- rep(1:kf,ceiling(n/kf))[1:n]
c3      <- data.table(k=t ,idx=sample(n))  
dt3     <- c3[,fitPred(raw[-idx,],raw[idx,],idx),by=.(k)]
kf.resid <- data.table(k=dt3$k,loc=dt3$idx,diff=dt3$resid)
if(demo){summary(kf.resid)}

```

Function to calculate parameterised and no parametrised confidence intervals

```{r}

ci <- function(res,a) {
  alpha <- a 
# Parametric             
  z     <- qnorm(1-(alpha/2) )
  t     <- mean(res)        
  s     <- sd(res)               
  lbP   <- t-z*s              
  ubP   <- t+z*s    
# Non parametric
  lo    <- alpha/2 # lower bound definition
  hi    <- 1-lo     # upper bound definition
  ci    <- quantile(res, c(lo,hi)) 
  results <- list(par_ci=data.table("lb"=lbP,"ub"=ubP),non_par_ci=ci)
return(results)
}

```

Confidence intervals of the different models

```{r}

ci(base.resid,0.05)
ci(cv.resid$diff,0.05)
ci(jk.resid$diff,0.05)
ci(kf.resid$diff,0.05)

```