---
title: "HW-5B"
output: html_document
date: "`r Sys.Date()`"
---

```{r}
rm(list=ls())
demo <- T
require(MASS)
data("airquality")

require(data.table)
require(tidyverse)
require(broom)

```

Adding airquality data into raw file and formatting the same

```{r}

raw     <- setDT(na.omit(airquality))[,-c("Day")]
raw$Month <- as.factor(raw$Month)
```

Establishing seed and number of iterations

```{r}
seed    <- 337337168

n      <- nrow(raw) 
b <- 500 
```

Setting lm function relation
```{r}
fmla <- Ozone ~ .
```

Classic bootstrapping 

```{r}
set.seed(seed)
t2 <- sample(n, size=b*n, replace=T) 
c2 <- data.table(g=rep(1:b,each=n),idx=t2)
cl.dt <- c2[,tidy(lm(fmla,data=raw[idx,]),conf.int=T), by=g]
if(demo) {summary(cl.dt)
print(cl.dt)}
```

Balanced bootstrapping

```{r}
set.seed(seed)
t3 <- sample(rep(1:n,b))
c3 <- data.table(g=rep(1:b,each=n),idx=t3)
bal.dt <- c3[,tidy(lm(fmla,data=raw[idx,]),conf.int=T),by=g] 
if(demo) {summary(bal.dt)
print(bal.dt)}
```

Function to obtain CIs and MSE
```{r}
ciPNP    <- function(v,ci) {  # data vector, 0 < CI < 1
         alpha <- 1-ci                 # area in the tails
         z     <- qnorm(1-alpha/2)     # Z for start of upper tail
         t     <- mean(v)              # for Paramteric
         s     <- sd(v)                # for Parametric
         e     <- mean(v-t) # Adding mse
         lP    <- round(t-z*s,digits = 3)              # lower bound for Parametric CI
         uP    <- round(t+z*s ,3)               # upper bound for Parametric CI
         bNP   <- c(alpha/2,1-alpha/2) # bounds for Non-Parametric CI
         ci    <- quantile(v, bNP)     # actual Non-Parametric CI bounds values
         return(data.table(avg=t,stDev=s,mse=e,loP=lP,     hiP=uP,
                                         loNP=round(ci[1],3), hiNP=round(ci[2],3)) )
         }
```

Classic bootstrap parameters

```{r}
coverage <- 0.95
ci2      <- cl.dt[,ciPNP(estimate,coverage),by=term]
print(ci2)
```

Balanced bootstrap
```{r}
ci3      <- bal.dt[,ciPNP(estimate,coverage),by=term]
print(ci3)
```

Baseline model

```{r}
m0 <- lm(fmla,data=raw)  
p0 <- m0$fitted.values  
r0 <- m0$residuals      
bs.dt <- as.data.table(tidy(m0,conf.int=T)) 
summary(m0)
```

Creating a function to get histogram plots of the distributions of the estimates

```{r}

plot_histogram <- function(df) {
  # loop through each row and create a histogram
  for (i in 1:nrow(df)) {
    hist(
      rnorm(10000, df$avg[i], df$stDev[i]),
      breaks = 20,
      main = df$term[i],
      xlab = "Value",
      ylab = "Frequency",
      xlim = c(df$loNP[i], df$hiP[i])
    )
    abline(v = df$hiP[i], col = "green")
    abline(v = df$loP[i], col = "green")
    abline(v = df$hiNP[i], col = "red")
    abline(v = df$loNP[i], col = "red")
    print(df[i])
  }
}

```


Classic bootstrap
```{r}
plot_histogram(ci2)
```

Balanced bootstrap
```{r}
plot_histogram(ci3)
```